---
title: "Week 1 Part 1"
output: pdf_document
---

This file consists of exercises for the course
*HarvardX PH525.1x Statistics and R* 

*Agnesh Panta*


*Exercise #2*
Create a numeric vector containing the numbers 2.23, 3.45, 1.87, 2.11, 7.33, 18.34, 19.23.

What is the average of these numbers?

```{r}
num_vec <- c(2.23, 3.45, 1.87, 2.11, 7.33, 18.34, 19.23)
mean(num_vec)
```
*Exercise #3*
Use a for loop to determine the value of $\sum_{i=1}^{25} i^2$

```{r}
sum_vec <- NULL

for (i in 1:25) {
  
  sum_vec[i] = i^2
  
  total_sum = sum(sum_vec)
  
}

total_sum
```

*Exercise #4*
The cars dataset is available in base R. You can type cars to see it. Use the class() function to determine what type of object is cars.
```{r}
class(cars)
```

*Exercise #5*

How many rows does the cars object have?
```{r}
nrow(cars)
```
*Exercise #6*

What is the name of the second column of cars?

```{r}
names(cars[2])
```
*Exercise #7*

The simplest way to extract the columns of a matrix or data.frame is using [. For example you can access the second column with cars[,2].

What is the average distance traveled in this dataset?

```{r}

dist_traveled <- data.frame(dist=cars[,2])

mean(dist_traveled[,1])

```
*Exercise #8*

Familiarize yourself with the which() function. Which row of cars has a a distance of 85?

```{r}
which(cars[,2]==85)
```

